---
title: STA610 Lab 6 Team 4 Report
author:
- Cole Juracek 
- Lauren Palazzo 
- Lingyu Zhou 
- Fan Zhu
date: "`r Sys.Date()`"
output:
  pdf_document: default
fontsize: 11pt
geometry: margin=2cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, fig.height = 3)
```

```{r}
df <- read.delim("BadgersFarmSurveysNoNA.txt", header = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(pacman)
pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, stringr, influence.ME, knitr, GGally, corrplot, correlation)
# devtools::install_github("goodekat/redres")
```
## Introduction 

Our data is drawn from the results of observational surveys of badger activities on 36 farms over three years. 
The aim of our analysis is to analyze factors associated with badger activity on farms, correlation over time of badger activity, and the level of variability in badger activity on the farms. 

## EDA

#### Response Variable: Signs_in_yard
```{r}
df$Signs_in_yard = factor(df$Signs_in_yard)
knitr::kable(table(df$Signs_in_yard), 
             col.names = c("Signs in Yard","Freq"), 
             caption = "Frequency Table for Signs in Yard")
```

#### Year & Season
```{r, fig.height=2, fig.align='center', include=FALSE}
df$Year = factor(df$Year)
df$Season = factor(df$Season)

bar_year <- ggplot(df, 
       aes(x = Year, 
           group = Signs_in_yard,
           fill = Signs_in_yard)) + 
  geom_bar(position = "dodge")

bar_season <- ggplot(df, 
       aes(x = Season, 
           group = Signs_in_yard,
           fill = Signs_in_yard)) + 
  geom_bar(position = "dodge")

grid.arrange(bar_year, bar_season, ncol=2)
```
```{r, fig.align='center', fig.height=2}
bar_year <- ggplot(df, 
       aes(x = Year, 
           fill = Signs_in_yard)) + 
  geom_bar(position = "fill") +
  labs(y = "Proportion")


bar_season <- ggplot(df, 
       aes(x = Season, 
           fill = Signs_in_yard)) + 
  geom_bar(position = "fill") +
  labs(y = "Proportion")

grid.arrange(bar_year, bar_season, ncol=2)
```

From the bar plot of Signs_in_yard by Year we can observe that the presence of badger activity in the farmyard in 2004 is lower than those in 2003 and 2005. From the bar plot of Signs_in_yard by Season we see that presence of badger activity is the highest in season 1 and the lowest in season 3. Even though badgers do not hibernate, they reduce activities during cold weathers so we will use a main effect of season in our model.

#### No_active_setts_in_fields & No_cattle_in_buidlings_yard
```{r, fig.height=2, fig.align='center', include=FALSE}
hist_No_active_setts_in_fields <- ggplot(df, aes(x=No_active_setts_in_fields)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 20)+
 geom_density(alpha=.2, fill="red") 

hist_No_cattle_in_buidlings_yard <- ggplot(df, aes(x=No_cattle_in_buidlings_yard)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 20)+
 geom_density(alpha=.2, fill="blue") 

grid.arrange(hist_No_active_setts_in_fields,hist_No_cattle_in_buidlings_yard, ncol=2)

# Both the distributions of No_active_setts_in_fields and No_cattle_in_buidlings_yard are extremely right skewed. 
```

```{r, fig.align='center', fig.height=2}
box_No_active_setts_in_fields <- ggplot(data=df, 
       mapping = aes(x = Signs_in_yard, 
                     y = No_active_setts_in_fields, 
                     fill = Signs_in_yard)) + 
  geom_boxplot()

box_No_cattle_in_buidlings_yard <- ggplot(data=df, 
       mapping = aes(x = Signs_in_yard, 
                     y = No_cattle_in_buidlings_yard, 
                     fill = Signs_in_yard)) + 
  geom_boxplot()

grid.arrange(box_No_active_setts_in_fields, 
             box_No_cattle_in_buidlings_yard, 
             ncol=2)
```
The median of No_active_setts_in_fields is much higher when there is badger activity in the farmyard than when there is no badger activity. The median of No_cattle_in_buidlings_yard is also much higher when there is badger activity in the farmyard than when there is no badger activity. 

#### Farm
```{r, fig.align='center', fig.height=2.5}
df$Accessible_feed_store_present = factor(df$Accessible_feed_store_present)

bar_farm <- ggplot(df, 
       aes(x = farm_code_numeric, 
           fill = Signs_in_yard)) + 
  geom_bar(position = "stack") + 
  labs(title = "Signs in Yard by Farm")

bar_Accessible_feed_store_present <- ggplot(df, 
       aes(x = Accessible_feed_store_present, 
           fill = Signs_in_yard)) + 
  geom_bar(position = "fill") +
  labs(y = "Proportion")

grid.arrange(bar_farm, bar_Accessible_feed_store_present, 
             ncol=2,
             widths=c(3, 2))
```
The patterns in the stacked bar chart of Signs_in_yard by farm indicate the incorporation of the random intercept by farm into our model. From the segmented bar chart for Accessible_feed_store_present we know that there are more presences of badger activities in the farmyard when there is Accessible_feed_store_present than when there is not. 

#### Binary Variables

```{r, fig.align='center', fig.height=3}
corrplot(cor(df[,c(13,14,15,16,19,20,21)]), 
         method = "color", 
         addCoef.col="grey", 
         order = "AOE",
         number.cex = 0.75,
         tl.cex = 0.5)
```

```{r, include=FALSE}
correlation::correlation(df[,c(10,13,14,15,16,19,20,21)])
```

```{r, include=FALSE}
table(df$Accessible_feed_store_present)
```

## Cole EDA (remaining variables)

```{r}
# For visualization, it helps to make these variables factors
df[, 11:21] <- apply(df[, 11:21], 2, factor)
```

First, we want to check the relative proportion of badger activity for each binary variable:

```{r fig.height=5}
plot_binary_data <- function(data, col) {
    var_df <- data %>% group_by_at(col) %>% summarise(prop = sum(Signs_in_yard == 1) / n())
    ggplot(var_df, aes_string(x=col, y='prop', fill=col)) + geom_col() + ylim(0, 1) + theme(legend.position = "none")
}

binary_plots <- lapply(names(df)[11:21], plot_binary_data, data=df)
grid.arrange(grobs=binary_plots, ncol = 4)
```

Nearly all of the proportions look like they *could* be similar - especially due to the small sample size. We can formally test whether the proportions are different with a Chi-square test:

```{r}
test_proportion <- function(col) {
    results <- chisq.test(table(df$Signs_in_yard, df[, col]))
    return(results$p.value)
}

good_ss_names <- c(11:17, 21)
results <- sapply(names(df)[good_ss_names], test_proportion)
kable(data.frame('p-val' = results), digits = 2)
```

At a significance level of $\alpha = 0.05$, we reject the null hypotheses for the following

- HayStraw
- Concentrates
- Molasses

And conclude their proportions are not equal.

For the remaining variables, the assumptions of a Chi-sq test were violated (Proteinblocks, Sugarbeet, and Vegetables). At least one of the table values has an expected counts < 5. We can test these with a Fisher exact test for small sample sizes:

```{r}
results <- sapply(c('Proteinblocks', 'Sugarbeet', 'Vegetables'), function(col) {
    results <- fisher.test(table(df$Signs_in_yard, df[, col]))
    return(results$p.value)
})
kable(data.frame('p-val' = results), digits = 3)
```

The results of the Fisher test suggest we should not reject the null at the standard significance level. We do not conclude a meaningful association between these variables and the response.

## Model
```{r}
mdl2 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric), 
              family = binomial(link="logit"),
              data = df)

mdl3 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season, 
              family = binomial(link="logit"),
              data = df)

mdl4 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season  + HayStraw  , 
              family = binomial(link="logit"),
              data = df)

mdl5 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season  + Sugarbeet, 
              family = binomial(link="logit"),
              data = df)

mdl6 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season  + Molasses, 
              family = binomial(link="logit"),
              data = df)

#anova(mdl4,mdl5,mdl6)


mdl7 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season  + Grass_silage, 
              family = binomial(link="logit"),
              data = df)

mdl8 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season + Cereal_silage, 
              family = binomial(link="logit"),
              data = df)

mdl9 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season + Cereal_grains, 
              family = binomial(link="logit"),
              data = df)

mdl10 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season  + Cereal_silage + Cereal_grains, 
              family = binomial(link="logit"),
              data = df)

#anova(mdl4, mdl7,mdl8,mdl9,mdl10)
```

```{r}
mdl_final <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season  + HayStraw , 
              family = binomial(link="logit"),
              data = df)
```

#### Model Selection
From the correlation plot we can observe that there are some significant correlations between the binary variables. Therefore, we need to exclude some of them to prevent the collinearity issue. For instance, Cereal_silage and HayStraw cannot exist in the same model because of their significant correlation coefficient. We did not use random intercept for season because there are only four seasons. Interaction terms are removed because they would result in failures to converge and singularity issues. Based on these principles, we built multiple viable models by adding combinations of binary variables on a basic model $Signs in yard \sim No active setts in fields+No cattle in buidlings yard+(1|farm code numeric)+Season$.

```{r}
BIC_res <- sapply(c(mdl4, mdl5, mdl6, mdl7, mdl8, mdl9, mdl10), BIC)
re_results <- data.frame('Model' = c('Model4', 
                                     'Model5', 
                                     'Model6', 
                                     'Model7',
                                     'Model8',
                                     'Model9',
                                     'Model10'),
                         'Binary Variables Added to the Basic Model' = c('HayStraw',
                                     'Sugarbeet',
                                     'Molasses',
                                     'Grass silage',
                                     'Cereal silage',
                                     'Cereal grains',
                                     'Cereal silage+Cereal grains'), 
                         'BIC' = BIC_res)
knitr::kable(re_results, caption = 'Model Comparisons with BIC', digits=2)
```

From the model comparisons above we chose model 4 as our final model following the principle of parsimony.

#### Model Specification
$$
y_{ij}|x_{ij} \sim Bernoulli(\pi_{ij})
$$
* $y_{ij}$ stands for response variable Signs_in_yard.
* i stands for individuals, i = 1,...,n. j stands for each farm, j = 1,...,36.

$$
\log(\frac{\pi_{ij}}{1-\pi_{ij}}) = \beta_0 + b_{0j} + \beta_{1}* x_1 + \beta_{2}*x_2 + \beta_{3}*x_3 + \beta_{4}*x_4
$$
$$
b_{0j} \sim N(0,\sigma^2)
$$

* $\beta_{0j}$ stands for random intercept by farm. 
* $x_1$ stands for number of active badger homes in nearby fields. (No_active_setts_in_fields).
* $x_2$ stands for number of cattle on the farm (No_cattle_in_buildings_yard).
* $x_3$ stands for Season.
* $x_4$ stands for HayStraw.

## Results
#### Fixed Effects
```{r, message=FALSE, warning=FALSE}
coefs <- data.frame(coef(summary(mdl_final)))
# use normal distribution to approximate p-value
coefs$Pr...z.. <- 2 * (1 - pnorm(abs(coefs$z.value)))
coefs$`Low CI Endpt` <- coefs$Estimate - 1.96*coefs$Std..Error
coefs$`High CI Endpt` <- coefs$Estimate + 1.96*coefs$Std..Error
knitr::kable(coefs[, c(1, 2, 5, 6)], col.names = c("Estimate", "Std.Error", "CI - Low", "CI - High"), caption = "Factors Related to Badger Activity", digits=2)
```

* *Intercept*: For a fixed farm (or across all farms), an observation, which has no number of active badger homes in nearby fields (no_active_setts_in_fields), no number of cattle on the farm (no_cattle_in_buildings_yard), Season being 1 and no HayStraw, has the odds of exp(-4.6049271) = 0.01 of badger activity being present in the farmyard.

* *Number of active badger homes in nearby fields (No_active_setts_in_fields)*: Controlling for other variables, for every one unit increase of number of active badger homes in nearby fields, the odds of badger activity being present in the farmyard increase by a multiplicative effect of exp(0.4878348) = 1.628786.

* *Number of cattle on the farm (No_cattle_in_buidlings_yard)*: Controlling for other variables, for every one unit increase of number of cattle on the farm, the odds of badger activity being present in the farmyard increase by a multiplicative effect of exp(0.0051664) = 1.00518.
  
* *Season* ("Season1" is the reference, other levels omitted for brevity):
    + Controlling for other variables, an observation in Season2 has exp(-0.2294778) = 0.7949486 times the odds of badger activity being present in the farmyard as an observation in Season1.
    
* *HayStraw* ("HayStraw0" is the reference):
    + Controlling for other variables, an observation with HayStraw1 has exp(-0.7805303) = 0.458163 times the odds of badger activity being present in the farmyard as an observation with HayStraw0.
    
* 95% confidence interval endpoints are also on the logit scale. For example, we can say that we are 95% confident that the odds estimate for HayStraw above is between exp(-1.9589250) = 0.14 and exp(0.3978645) = 1.49. 

#### Random Effects
```{r}
re_dat = as.data.frame(VarCorr(mdl_final))
re_names = rbind("Farm (Intercept)")
re_results <- cbind(re_names, re_dat$vcov, re_dat$sdcor)
knitr::kable(re_results, col.names = c("Group", "Variance", "Std.Dev."), digits = 2, caption = "Farm Variation")
```


```{r, include=FALSE}
library(sjstats)
icc(mdl_final)
#Adjusted ICC: 0.601
#Conditional ICC: 0.484
```

The variance of the farm-level intercepts (i.e., the baseline log-odds of badger activity among the farms) is about 4.94, while the grand intercept is centered at about -4.60. In other words, the variance is about the same as the mean, which suggests a considerable amount of between-farm heterogeneity.

The adjusted ICC is 0.601, which indicates high similarity between values from the same farm. I.E., each farm is fairly homogeneous with respect to its observations of badger activity. 

```{r, fig.align='center', fig.height=4.5}
dotplot(ranef(mdl_final, condVar=TRUE))$farm_code_numeric
```

The odds of the random intercepts range from exp(4.29665538) = 73.45371 for farm with code 3 to exp(-2.39464632) = 0.09120493 for farm with code 8. We interpret this as follows: the estimated odds of badger activity being present in the farmyard is highest for farm with code 3 with zero active badger homes in nearby fields (no_active_setts_in_fields), zero cattle on the farm (no_cattle_in_buildings_yard), Season being 1 and no HayStraw. One interesting observation from the above graph is that farms with below-average intercepts appear to be generally somewhat more homogeneous (similar in value) than those with above-average intercepts. 

With regard to correlation over time, the correlations of fixed effects in the model summary suggest how the coefficients of 'Season' tend to correlate with each other. For example, Season 1's respective correlations with Seasons 2, 3, and 4 are -0.444, -0.463, and -0.232. So, higher badger activity in Season 1 is expected to be associated with lower badger activity in the other three seasons. 

```{r include=FALSE}
summary(mdl_final)
```

\newpage
## Appendix

```{r, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
