---
title: \vspace{-1.5cm} \textbf{STA610 Lab 6 Team 4 Report}
author:
- Cole Juracek 
- Lauren Palazzo 
- Lingyu Zhou 
- Fan Zhu
date: "`r Sys.Date()`"
output:
  pdf_document: default
fontsize: 11pt
geometry: margin=2cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, fig.height = 3)
```

```{r}
df <- read.delim("BadgersFarmSurveysNoNA.txt", header = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(pacman)
pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres, stringr, influence.ME, knitr, GGally, corrplot)
# devtools::install_github("goodekat/redres")
```

## EDA
#### Response Variable: Signs_in_yard
```{r}
df$Signs_in_yard = factor(df$Signs_in_yard)
knitr::kable(table(df$Signs_in_yard), 
             col.names = c("Signs in Yard","Freq"), 
             caption = "Frequency Table for Signs in Yard")
```

#### Year & Season
```{r, fig.height=2, fig.align='center', include=FALSE}
df$Year = factor(df$Year)
df$Season = factor(df$Season)

bar_year <- ggplot(df, 
       aes(x = Year, 
           group = Signs_in_yard,
           fill = Signs_in_yard)) + 
  geom_bar(position = "dodge")

bar_season <- ggplot(df, 
       aes(x = Season, 
           group = Signs_in_yard,
           fill = Signs_in_yard)) + 
  geom_bar(position = "dodge")

grid.arrange(bar_year, bar_season, ncol=2)
```
```{r, fig.align='center', fig.height=2}
bar_year <- ggplot(df, 
       aes(x = Year, 
           fill = Signs_in_yard)) + 
  geom_bar(position = "fill") +
  labs(y = "Proportion")


bar_season <- ggplot(df, 
       aes(x = Season, 
           fill = Signs_in_yard)) + 
  geom_bar(position = "fill") +
  labs(y = "Proportion")

grid.arrange(bar_year, bar_season, ncol=2)
```

From the bar plot of Signs_in_yard by Year we can observe that the presence of badger activity in the farmyard in 2004 is lower than those in 2003 and 2005. From the bar plot of Signs_in_yard by Season we see that presence of badger activity is the highest in season 1 and the lowest in season 3. Even though badgers do not hibernate, they reduce activities during cold weathers so we will use a main effect of season in our model.

#### No_active_setts_in_fields & No_cattle_in_buidlings_yard
```{r, fig.height=2, fig.align='center', include=FALSE}
hist_No_active_setts_in_fields <- ggplot(df, aes(x=No_active_setts_in_fields)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 20)+
 geom_density(alpha=.2, fill="red") 

hist_No_cattle_in_buidlings_yard <- ggplot(df, aes(x=No_cattle_in_buidlings_yard)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 20)+
 geom_density(alpha=.2, fill="blue") 

grid.arrange(hist_No_active_setts_in_fields,hist_No_cattle_in_buidlings_yard, ncol=2)

# Both the distributions of No_active_setts_in_fields and No_cattle_in_buidlings_yard are extremely right skewed. 
```

```{r, fig.align='center', fig.height=2}
box_No_active_setts_in_fields <- ggplot(data=df, 
       mapping = aes(x = Signs_in_yard, 
                     y = No_active_setts_in_fields, 
                     fill = Signs_in_yard)) + 
  geom_boxplot()

box_No_cattle_in_buidlings_yard <- ggplot(data=df, 
       mapping = aes(x = Signs_in_yard, 
                     y = No_cattle_in_buidlings_yard, 
                     fill = Signs_in_yard)) + 
  geom_boxplot()

grid.arrange(box_No_active_setts_in_fields, 
             box_No_cattle_in_buidlings_yard, 
             ncol=2)
```
The median of No_active_setts_in_fields is much higher when there is badger activity in the farmyard than when there is no badger activity. The median of No_cattle_in_buidlings_yard is also much higher when there is badger activity in the farmyard than when there is no badger activity. 

#### Farm
```{r, fig.align='center', fig.height=2.5}
df$Accessible_feed_store_present = factor(df$Accessible_feed_store_present)

bar_farm <- ggplot(df, 
       aes(x = farm_code_numeric, 
           fill = Signs_in_yard)) + 
  geom_bar(position = "stack") + 
  labs(title = "Signs in Yard by Farm")

bar_Accessible_feed_store_present <- ggplot(df, 
       aes(x = Accessible_feed_store_present, 
           fill = Signs_in_yard)) + 
  geom_bar(position = "fill") +
  labs(y = "Proportion")

grid.arrange(bar_farm, bar_Accessible_feed_store_present, 
             ncol=2,
             widths=c(3, 2))
```
The patterns in the stacked bar chart of Signs_in_yard by farm indicate the incorporation of the random intercept by farm into our model. From the segmented bar chart for Accessible_feed_store_present we know that there are more presences of badger activities in the farmyard when there is Accessible_feed_store_present than when there is not. 

#### Binary Variables

```{r, fig.align='center', fig.height=3}
corrplot(cor(df[,c(13,14,15,16,19,20,21)]), 
         method = "color", 
         addCoef.col="grey", 
         order = "AOE",
         number.cex = 0.75,
         tl.cex = 0.5)
```


```{r, include=FALSE}
correlation::correlation(df[,c(10,13,14,15,16,19,20,21)])
```


```{r, include=FALSE}
table(df$Accessible_feed_store_present)
```




## Model
```{r}
mdl2 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric), 
              family = binomial(link="logit"),
              data = df)

mdl3 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season, 
              family = binomial(link="logit"),
              data = df)

mdl4 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season  + HayStraw  , 
              family = binomial(link="logit"),
              data = df)

mdl5 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season  + Sugarbeet, 
              family = binomial(link="logit"),
              data = df)

mdl6 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season  + Molasses, 
              family = binomial(link="logit"),
              data = df)

#anova(mdl4,mdl5,mdl6)


mdl7 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season  + Grass_silage, 
              family = binomial(link="logit"),
              data = df)

mdl8 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season + Cereal_silage, 
              family = binomial(link="logit"),
              data = df)

mdl9 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season + Cereal_grains, 
              family = binomial(link="logit"),
              data = df)

mdl10 <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season  + Cereal_silage + Cereal_grains, 
              family = binomial(link="logit"),
              data = df)

#anova(mdl4, mdl7,mdl8,mdl9,mdl10)
```

```{r}
mdl_final <- glmer(Signs_in_yard ~ No_active_setts_in_fields + No_cattle_in_buidlings_yard + (1|farm_code_numeric) + Season  + HayStraw , 
              family = binomial(link="logit"),
              data = df)
```

#### Model Selection
From the correlation plot we can observe that there are some significant correlations between the binary variables. Therefore, we need to exclude some of them to prevent the collinearity issue. For instance, Cereal_silage and HayStraw cannot exist in the same model because of their significant correlation coefficient. We did not use random intercept for season because there are only four seasons. Interaction terms are removed because they would result in failures to converge and singularity issues. Based on these principles, we built multiple viable models by adding combinations of binary variables on a basic model $Signs in yard \sim No active setts in fields+No cattle in buidlings yard+(1|farm code numeric)+Season$.

```{r}
BIC_res <- sapply(c(mdl4, mdl5, mdl6, mdl7, mdl8, mdl9, mdl10), BIC)
re_results <- data.frame('Model' = c('Model4', 
                                     'Model5', 
                                     'Model6', 
                                     'Model7',
                                     'Model8',
                                     'Model9',
                                     'Model10'),
                         'Binary Variables Added to the Basic Model' = c('HayStraw',
                                     'Sugarbeet',
                                     'Molasses',
                                     'Grass silage',
                                     'Cereal silage',
                                     'Cereal grains',
                                     'Cereal silage+Cereal grains'), 
                         'BIC' = BIC_res)
knitr::kable(re_results, caption = 'Model Comparisons with BIC')
```

From the model comparisons above we chose model 4 as our final model following the principle of parsimony.

#### Model Specification


## Result
#### Fixed Effects
```{r, message=FALSE, warning=FALSE}
coefs <- data.frame(coef(summary(mdl_final)))
# use normal distribution to approximate p-value
coefs$Pr...z.. <- 2 * (1 - pnorm(abs(coefs$z.value)))
knitr::kable(coefs, col.names = c("Estimate", "Std.Error", "t-value", "P-value"))
```

#### Random Effects
```{r}
re_dat = as.data.frame(VarCorr(mdl_final))
re_names = rbind("Farm (Intercept)")
re_results <- cbind(re_names, re_dat$vcov, re_dat$sdcor)
knitr::kable(re_results, col.names = c("Group", "Variance", "Std.Dev."), digits = 4)
```

```{r, fig.align='center', fig.height=6}
dotplot(ranef(mdl_final, condVar=TRUE))$farm_code_numeric
```

```{r, include=FALSE}
library(sjstats)
icc(mdl_final)
#Adjusted ICC: 0.601
#Conditional ICC: 0.484
```
